---
title: "RamirezJ,Hmwk6"
author: "Josie Ramirez"
date: "2024-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Downloading Data
Focusing on getting the needed variables for one year, 2023. 
Once I am able to do one year, I can do the same for the other seven years 
```{r}
getwd()
setwd("C:/Users/Ramirez/OneDrive/Desktop/Cla$$/Fall '24/ds340h/atusRQ")

act<-read.csv("atusact_2023.dat",header=TRUE)
cpus<- read.csv("atuscps_2023.dat", header = TRUE)

head(act)
head(cpus)

```

## Working with Activity 2023
```{r}
library(tidyverse)
#install.packages("lubridate")
#install.packages("hms")

#library(lubridate)
library(hms)



#trying to use tidyverse rather than r [] syntax

```


```{r}

#variables wanted: TUCASEID, TESEX, TUACTIVITY_N, TUSTARTTIM, TUSTOPTIME, TUTIER1CODE, TUTIER2CODE, TUTIER3CODE

newAct<- act %>%
  filter((TUTIER1CODE == 03 | TUTIER1CODE == 04) &
           (TUTIER2CODE == 1 | TUTIER2CODE == 2 | TUTIER2CODE == 3)
           ) %>%
  select(TUCASEID, TUACTIVITY_N, TUSTARTTIM, TUSTOPTIME, TUTIER1CODE, TUTIER2CODE, TUTIER3CODE)

#TESEX not in activity would need to see another file
#---ask: not sure if I should include this as an analysis

head(newAct)
dim(newAct)

```
## Focusing on getting total time for Household Children


```{r}
class(newAct$TUSTARTTIM)
#need to convert start and stop times 
```

```{r}
hh <- newAct[newAct$TUTIER1CODE == 3, ]
dim(hh)
#these indicate specifically household children care

hh<- hh %>%
  mutate(total_mins = as.numeric(difftime(as_hms(TUSTOPTIME), as_hms(TUSTARTTIM), units = "mins")) )

head(hh)

```
## For household children, this is the total amount of time spent
```{r}
minsBy2Code <- aggregate(total_mins ~ TUTIER2CODE, data = hh, FUN = sum)
minsBy2Code

```

## Focusing on getting total time for Non Household Children

```{r}
nonhh <- newAct[newAct$TUTIER1CODE == 4, ]
dim(hh)
#these indicate specifically household children care

nonhh<- nonhh %>%
  mutate(total_mins = as.numeric(difftime(as_hms(TUSTOPTIME), as_hms(TUSTARTTIM), units = "mins")) )

head(nonhh)

minsBy2Code <- aggregate(total_mins ~ TUTIER2CODE, data = nonhh, FUN = sum)
minsBy2Code
```

## Combing datasets so info is all in one
```{r}
act23 <- full_join(hh, nonhh, by = "TUCASEID")
head(act23)
tail(act23)
#not really sure
#-------------ask
```


