---
title: "RamirezJ,Hmwk6"
author: "Josie Ramirez"
date: "2024-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Downloading Data

Focusing on getting the needed variables for one year, 2023. Once I am able to do one year, I can do the same for the other seven years

```{r}
getwd()
#setwd("C:/Users/Ramirez/OneDrive/Desktop/Cla$$/Fall '24/ds340h/atusRQ")
#C:\Users\Ramirez\OneDrive\Desktop\ds340hcapstone\atusRQ

#setwd("C:/Users/Ramirez/OneDrive/Desktop/ds340hcapstone/atusRQ")

# act<-read.csv("atusact_2023.dat",header=TRUE)
# cpus<- read.csv("atuscps_2023.dat", header = TRUE)
# 
# head(act)
# head(cpus)

```

## Working with Activity 2023

```{r}
library(tidyverse)
#install.packages("lubridate")
#install.packages("hms")

#library(lubridate)
library(hms)

#library(ggplot2) #already loaded with tidyverse



#trying to use tidyverse rather than r [] syntax

```

```{r}

#variables wanted: TUCASEID, TESEX, TUACTIVITY_N, TUSTARTTIM, TUSTOPTIME, TUTIER1CODE, TUTIER2CODE, TUTIER3CODE

# newAct<- act %>%
#   filter((TUTIER1CODE == 03 | TUTIER1CODE == 04) &
#            (TUTIER2CODE == 1 | TUTIER2CODE == 2 | TUTIER2CODE == 3)
#            ) %>%
#   select(TUCASEID, TUACTIVITY_N, TUSTARTTIM, TUSTOPTIME, TUTIER1CODE, TUTIER2CODE, TUTIER3CODE)
# 
# #TESEX not in activity would need to see another file
# 
# head(newAct)
# dim(newAct)

```

```{r}
# newAct<- newAct %>%
#     mutate(total_mins = as.numeric(difftime(as_hms(TUSTOPTIME), as_hms(TUSTARTTIM), units = "mins")) ) %>%
#     mutate(year = 2023)
# 
# newAct
```

## Creating Function to merge calculate total time for all 8 years

```{r}

#loading in all the data sets

act23<-read.csv("atusact_2023.dat",header=TRUE)
act22<-read.csv("atusact_2022.dat",header=TRUE)
act21<-read.csv("atusact_2021.dat",header=TRUE)
act20<-read.csv("atusact_2020.dat",header=TRUE)
act19<-read.csv("atusact_2019.dat",header=TRUE)
act18<-read.csv("atusact_2018.dat",header=TRUE)
act17<-read.csv("atusact_2017.dat",header=TRUE)
act16<-read.csv("atusact_2016.dat",header=TRUE)



#making them into a list
myList = list()
myList[["df1"]] <- act23
myList[["df2"]] <- act22
myList[["df3"]] <- act21
myList[["df4"]] <- act20
myList[["df5"]] <- act19
myList[["df6"]] <- act18
myList[["df7"]] <- act17
myList[["df8"]] <- act16

#print(names(myList))

#lapply(myList, head)

#loop through every element in that list
y = 2023

dfList = list()

for (i in myList){
  actY<- i %>%
  mutate(total_mins = as.numeric(difftime(as_hms(TUSTOPTIME), as_hms(TUSTARTTIM), units = "mins")) ) %>%
  mutate(year = y)

  ##Taking only the columns I want

  
  newActY<- actY %>%
  filter((TUTIER1CODE == 3 | TUTIER1CODE == 4) &
           (TUTIER2CODE == 1 | TUTIER2CODE == 2 | TUTIER2CODE == 3)
           ) %>%
  select(TUCASEID, TUACTIVITY_N, TUSTARTTIM, TUSTOPTIME, TUTIER1CODE, TUTIER2CODE, TUTIER3CODE, total_mins, year)
  
  dfList[[paste0("year", y)]] <- newActY
  #myList[[as.character(y)]] <- newActY
  
  
  y<- y-1
}
  
print(names(dfList))

lapply(dfList, head)


```

## Merging all the datasets, rbind

```{r}

full_df<- rbind(dfList[["year2023"]], dfList[["year2022"]], 
                dfList[["year2021"]], dfList[["year2020"]], 
                dfList[["year2019"]], dfList[["year2018"]], 
                dfList[["year2017"]], dfList[["year2016"]])

head(full_df)
tail(full_df)


```

## Total Time for Household and NonHousehold Children

```{r}
time<- full_df %>%
  group_by(TUTIER1CODE) %>%
  summarise(mean_time = mean(total_mins), sum_mins = sum(total_mins), sum_hr = sum(total_mins)/60)
time

```

## Household Children: Total Time Spent by Tier2Code

```{r}
timeHH_type<- full_df %>%
  filter(TUTIER1CODE==3)%>%
  group_by(TUTIER2CODE) %>%
  summarise(mean_time = mean(total_mins), sum_mins = sum(total_mins), sum_hr = sum(total_mins)/60)

timeHH_type

```

## NonHousehold Children: Total Time Spent by Tier2Code

```{r}
timeNH_type<- full_df %>%
  filter(TUTIER1CODE==4)%>%
  group_by(TUTIER2CODE) %>%
  summarise(mean_time = mean(total_mins), sum_mins = sum(total_mins), sum_hr = sum(total_mins)/60)

timeNH_type
```

## Visual showing difference between total time spent between household and nonhousehold children

```{r}
full_df

ggplot(full_df, aes(x = factor(TUTIER1CODE), y = total_mins)) +
         geom_violin(trim = FALSE) +
         labs(x = "Children", y = "Total Time Spent with Child", title = paste("In the past 8 years, total time spent caring for and helping","\nhousehold and nonhousehold children, capped at 250")) + theme_minimal() +
    scale_x_discrete(labels = c("Household Children", "Nonhousehold Children")) +  # Custom labels +
  
  # Add mean and median
  stat_summary(aes(color = "Mean"), fun = mean, geom = "point",size = 1) +  
  stat_summary(aes(color = "Median"), fun = median, geom = "point", size = 1) +

  scale_color_manual(values = c("Mean" = "blue", "Median" = "yellow")) +  # Custom colors for legend
  guides(color = guide_legend(title = NULL))  +

  scale_y_continuous(limits = c(0, 250), breaks = seq(0, 650, by = 50)) +
  theme(plot.title = element_text(hjust = 0.5))
  

#maybe would want to zoom in more
#the x-axis could be better




```

Left off here: \#################################################################

#Trying to Merge Activity with Income

```{r}
head(newAct)
head(cpus)

#i want to add the income variable for all tucaseid in newAct, other values drop

mergedDf<- newAct %>%
  left_join(select(cpus, TUCASEID, HEFAMINC), by = "TUCASEID")

mergedDf

#realized i also need the total respondents to be able to calculate the federal measurement code

```

## Code to be ignored:

```{r}
#############
# code I'm not using anymore 
#############






# grouped23<- newAct %>%
#   group_by(TUTIER1CODE) %>%
#   summarise(mean_time = mean(total_mins), sum_mins = sum(total_mins), sum_hr = sum(total_mins)/60)
# grouped23


# grouped23_type<- newAct %>%
#   filter(TUTIER1CODE==3)%>%
#   group_by(TUTIER2CODE) %>%
#   summarise(mean_time = mean(total_mins), sum_mins = sum(total_mins), sum_hr = sum(total_mins)/60)
# 
# grouped23_type





## Focusing on getting total time for Household Children



# class(newAct$TUSTARTTIM)
# #need to convert start and stop times 
# 
# 
# 
# hh <- newAct[newAct$TUTIER1CODE == 3, ]
# dim(hh)
# #these indicate specifically household children care
# 
# hh<- hh %>%
#   mutate(total_mins = as.numeric(difftime(as_hms(TUSTOPTIME), as_hms(TUSTARTTIM), units = "mins")) )
# 
# dim(hh)
# head(hh)
# 
# 
# ## For household children, this is the total amount of time spent
# 
# minsBy2Code <- aggregate(total_mins ~ TUTIER2CODE, data = hh, FUN = sum)
# minsBy2Code
# 
# 
# 
# ## Focusing on getting total time for Non Household Children
# 
# nonhh <- newAct[newAct$TUTIER1CODE == 4, ]
# dim(hh)
# #these indicate specifically household children care
# 
# nonhh<- nonhh %>%
#   mutate(total_mins = as.numeric(difftime(as_hms(TUSTOPTIME), as_hms(TUSTARTTIM), units = "mins")) )
# 
# head(nonhh)
# dim(nonhh)
# 
# minsBy2Code <- aggregate(total_mins ~ TUTIER2CODE, data = nonhh, FUN = sum)
# minsBy2Code
# 
# ## Combing datasets so info is all in one
# act23 <- merge(hh, nonhh, by = "TUCASEID", all = FALSE)
# head(act23)
# tail(act23)
# #not really sure
# #-------------ask

```
